---
title: "Lab4"
author: "Sparsh Raj_19BPS1028"
date: "19/02/2022"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


# 1. Import dataset and display it and check for unfilled data.
```{r}
library(tseries)
library(forecast)

data<-AirPassengers
data

sum(is.na(data))
```


# 2. Display the statistical info of the dataset such as min, max, 1st quartile, 3rd quartile, mean and median.
```{r}
dMin<-min(data)
print(paste("Min = ",dMin))

dMax<-max(data)
print(paste("Max = ",dMax))

dMean<-mean(data)
print(paste("Mean = ",dMean))

dMed<-median(data)
print(paste("Median = ",dMed))

quan<-quantile(data)
print("Quartiles are = ")
quan
```

# 3. Plot data
```{r}
plot(data,xlab="Date", ylab = "Passenger numbers (1000's)",main="Air Passenger numbers from 1949 to 1961")
```

# 4. Plot as timeseries ‘data’ (monthwise)

```{r}
boxplot(data~cycle(data),xlab="Date", ylab = "Passenger Numbers (1000's)" ,main ="Monthly Air Passengers Boxplot from 1949 to 1961")
```


# 5. Decompose the data as multiplicative and store as ‘ddata’
```{r}
ddata <- decompose(data,"multiplicative")
```


# 6. & 7. Plot ddata and also plot the following: trend, seasonal and random.
```{r}
autoplot(ddata)
```


# 8. Perform ADF test for stationarity.
```{r}
adf.test(data) 
```
 

# 9. Plot ACF and PACF


Correlogram of Air Passengers from 1949 to 1961
```{r}
autoplot(acf(data,plot=FALSE)) 
```

Review random time series for any missing values
```{r}
ddata$random
```


Correlogram of Air Passengers Random Component from 1949 to 1961


Autoplot the random time series from 7:138 which exclude the NA values
```{r}
autoplot(acf(ddata$random[7:138],plot=FALSE)) 
```



# 10. Model using ARIMA
```{r}
fitData <- auto.arima(data)
fitData

library(ggfortify)
ggtsdiag(fitData)
```

